Mammalian protein-coding genes exhibit widespread beneficial mutations that are not adaptive

### 2023-04-03 manuscript/main.tex
Updated manuscript before NG corrections

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\der}{\text{d}}
\newcommand{\e}{\text{e}}
\newcommand{\Ne}{N_{\text{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\text{fix}}}
\newcommand{\dn}{d_N}
\newcommand{\ds}{d_S}
\newcommand{\dnds}{\dn / \ds}
\newcommand{\Sphy}{S_{0}}
\newcommand{\Sphyclass}{\mathcal{C}}
\newcommand{\given}{\mid}
\newcommand{\Spop}{S}
\newcommand{\PpolyDel}{\proba [ \Spop^{\text{d}}]}
\newcommand{\PpolyNeutral}{\proba [ \Spop^{\text{n}} ]}
\newcommand{\PpolyAdv}{\proba [ \Spop^{\text{b}} ]}
\newcommand{\AdvMean}{\beta_b}
\newcommand{\DelMean}{\beta_d}
\renewcommand{\baselinestretch}{1.5}
\renewcommand{\arraystretch}{0.6}
\linenumbers
\frenchspacing

@@

    Adaptation is considered one of the main processes shaping the diversity of forms and functions across the tree of life\cite{darwin_origin_1859}.
    In evolution, adaptation is tightly linked to the notion of environmental change and to species responding to this change\cite{merrell_adaptive_1994, gavrilets_adaptive_2009}.
    For adaptation to take place, there needs to be variation within populations, which at the molecular level mostly appears via mutations in the DNA sequence.
    While neutral mutations will have no impact on the fitness of an individual, deleterious mutations have a negative effect and beneficial mutations lead their bearer toward a higher fitness.
    Compared to a neutral mutation, a beneficial mutation is thus more likely to reach fixation in the population, resulting in a substitution at the species level.
    Upon environmental change, adaptive beneficial mutations toward new fitness optima are more likely to rise in frequency, resulting in an increased substitution rate (fig.~\ref{fig:fitness-landscape}A).
    Altogether, an accelerated substitution rate is thus commonly interpreted as a sign of adaptation\cite{mcdonald_adaptative_1991, smith_adaptive_2002, welch_estimating_2006}.
    The current availability of large-scale genomic data and the development of theoretical models have enabled us to detect and quantify acceleration of substitution rates across genes and lineages, and these analyses have become common practice in evolutionary biology\cite{yang_statistical_2000, eyre-walker_genomic_2006, moutinho_variation_2019}.
    These approaches have led to a better understanding of the processes determining the rates of molecular evolution.
    However, they have had the collateral effect of associating beneficial mutations with adaptive evolution when, in fact, adaptive evolution is not the only process that can lead to beneficial mutations\cite{charlesworth_other_2007, mustonen_fitness_2009}.

    In a constant environment, a deleterious mutation can reach fixation in the population by genetic drift\cite{Ohta1992}.
    Subsequently, a new mutation restoring the ancestral fitness will thus be beneficial (fig.~\ref{fig:fitness-landscape}B), even though no change in the environment has occurred\cite{hartl_compensatory_1996, sella_application_2005, mustonen_fitness_2009, cvijovic_fate_2015}.
    The restoration of the ancestral fitness can happen with a mutation occurring either at a different locus from that of the initial mutation, in which case it is referred to as a compensatory mutation\cite{hartl_compensatory_1996, mustonen_fitness_2009}, or at the same locus of the initial mutation, referred to as a beneficial back-mutation\cite{piganeau_estimating_2003, charlesworth_other_2007}.
    Now, while compensatory mutations induce genetic changes at the sequence level, and thus genetic diversification, beneficial back-mutations reduce the inter-specific genetic diversity and do not contribute to genetic innovation.
    Although beneficial back-mutations were considered negligible in the enunciation of the nearly-neutral theory by Tomoko Ohta\cite{Ohta1992}, their importance during a population size expansion has been acknowledged more recently\cite{charlesworth_other_2007}.
    However, differentiating between an adaptive mutation and a beneficial back-mutation has remained challenging\cite{chi_detecting_2020}.
    In this study, we seek to find empirical evidence for beneficial back-mutations and to test long-standing questions in molecular evolution such as how common are beneficial back-mutations, and
    what is the probability for any beneficial mutation observed in current populations to be a back-mutation.
    By addressing these questions we provide novel insight on the process of adaptation and, more broadly, on how natural selection has affected protein-coding DNA sequences throughout evolution.

    \subsection*{Fitness landscape reconstruction}

    An adaptive mutation responding to a change in the environment or a beneficial back-mutation have equivalent consequences for their bearer\cite{charlesworth_other_2007}.
    Similarly, at the level of the population, both types of mutations will result in a positive transmission bias of the beneficial allele.
    However, at the macro-evolutionary scale the consequences of these two types of mutations are fundamentally different.
    While adaptive mutations promote phenotype diversification (fig.~\ref{fig:fitness-landscape}C), beneficial back-mutations promote phenotype stability and may help preserve well established biological systems (fig.~\ref{fig:fitness-landscape}D).
    Additionally, the direction of adaptive evolution is unpredictable because it is caused by an unpredictable change in the environment and, hence, in the underlying fitness landscape\cite{bazykin_changing_2015}.
    On the other hand, beneficial back-mutations are predictable if the underlying stable fitness landscape is known because changes from non-optimal amino acids to optimal ones will be expected.
    In this context, the mutation-selection framework permits to relate the patterns of substitution along a phylogeny to the underlying fitness landscape\cite{halpern_evolutionary_1998, mccandlish_modeling_2014}.
    Such mutation-selection models applied to protein coding DNA sequences allow us to estimate relative fitnesses for all amino-acids for each site of the sequence, explicitly assuming that the underlying fitness landscape is stable along the phylogenetic tree\cite{rodrigue_mechanistic_2010, tamuri_estimating_2012, rodrigue_detecting_2017}.
    Importantly, because mutation-selection models at the phylogenetic scale are based on population-genetics equations, their estimates of fitness effects are directly interpretable as fitness effects for an individual within a population.
    %These methods account for phylogenetic relationships, the structure of the genetic code and accommodate for underlying mutational biases\cite{rodrigue_mechanistic_2010, tamuri_estimating_2012, latrille_improved_2022}.

    Theoretically, we can thus use large-scale genomic data to assess whether the fitnesses estimated at the phylogenetic scale are predictive of the fitness effect at the population scale, for both deleterious and beneficial back-mutations.
    The placental mammals represent an excellent study system to perform such analysis.
    Having originated $\sim$100 Mya\cite{kumar_timetree_2017}, they diversified quickly.
    Additionally, polymorphism data are available for many species\cite{howe_ensembl_2021} as high quality protein-coding DNA alignments across the genome\cite{ranwez_orthomam_2007, scornavacca_orthomam_2019}.
    By restricting our analysis to protein-coding orthologous genes across all the phylogeny, we restrict our analysis to genes mostly under functional constrains across mammals.
    For each gene, fitting the mutation-selection model to the multi-species sequence alignment assuming that the underlying fitness landscape is stable along the phylogenetic tree allows obtain relative fitnesses for all amino-acids for each site of the alignment (fig~\ref{fig:method}A).
    From the fitness landscape, we can calculate the scaled selection coefficient $\Sphy$ for any possible amino-acid change at each position.
    From the selection coefficient, we can classify any mutation as either a deleterious mutation toward a less fit amino-acid ($\Sphy^{\text{d}}$), a nearly-neutral mutation ($\Sphy^{\text{n}}$) or a beneficial back-mutation toward a known fitter amino-acid ($\Sphy^{\text{b}}$).
    Having identified which potential DNA changes represent beneficial back-mutations (fig~\ref{fig:method}B), we retrieved polymorphism data from 28 populations belonging to 6 genera (\textit{Equus}, \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus} and \textit{Homo}) to assess the presence of beneficial back-mutations at the population scale.
    We focused on currently segregating mutations within populations and also on substitutions in the terminal branches, and checked if any of these observed changes where beneficial back-mutations (fig~\ref{fig:method}C-E).
    Essentially, by integrating large scale genomic datasets at both phylogenetic and population scales we have been able to estimate the amount of beneficial back-mutations across the entire exome for 6 different mammalian genera (fig~\ref{fig:method}F-H).

@@

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.method.proba}
        \caption{
            Selection coefficients at the phylogenetic and population scales.
            At the phylogenetic scale (A), the amino-acid Wrightian fitness for each site is estimated from protein-coding DNA alignments using mutation-selection codon models.
            For every possible mutation, the difference in amino-acid fitness before and after the mutation allows us to compute the selection coefficient at the phylogenetic scale ($\Sphy$).
            Mutations with $\Sphy < -1$ are considered deleterious ($\Sphy^{\text{d}}$ in red), and mutations with $-1 < \Sphy < 1$ are considered nearly-neutral ($\Sphy^{\text{n}}$ in green).
            Mutations with $\Sphy > 1$ are toward a fitter amino-acids, and thus are considered beneficial back-mutations ($\Sphy^{\text{b}}$ in green blue), repairing existing functions (B).
            At the population scale (C), each observed single nucleotide polymorphism (SNP) segregating in the population is classified according to its $\Sphy$ value.
            Similarly, substitutions along the terminal branch are classified depending on their $\Sphy$ value.
            Together, substitutions and polymorphisms are contrasted to synonymous changes used as a neutral expectation such as estimate selection coefficients at the population scale ($\Spop$).
            For each class of selection ($\Sphy^{\text{b}}$, $\Sphy^{\text{n}}$, $\Sphy^{\text{d}}$), we can thus assess whether $\Sphy$ predicts $\Spop$ (E), and compute \textit{precision} (F-G) and \textit{recall} (F-H) for each class.
            The \textit{recall} value for the class $\Sphy^{\text{b}}$ is the probability of back-mutations among all beneficial ones (H).
        }
        \label{fig:method}
    \end{figure*}

    \subsection*{Selection along the terminal branches}
    First, we assessed whether fitness effects derived from the mutation-selection model at the phylogenetic scale are predictive of selection occurring in terminal branches.
    To achieve this, we recovered the mutations that reached fixation in the terminal branches of 6 different mammalian genera.
    Only mutations fixed in a population were considered as substitutions in the corresponding branch, while segregating mutations were discarded from this analysis.
    For each substitution identified in the terminal branches, we obtained its scaled selection coefficient ($\Sphy$) as the change in amino-acid fitness at the corresponding codon site in the mammalian alignment (fig.~\ref{fig:method}A-C).
    Importantly, the mammalian alignment used to estimate amino-acid fitness landscape does not include the genera for which we reconstructed the substitutions, to avoid fallacy and circularity in the reasoning.
    Because $\Sphy$ was based on estimation of fitness effects at the mammalian scale, mutations with $\Sphy>1$ were considered beneficial back-mutations ($\Sphy^{\text{b}}$), toward a fitter amino acid.
    Among all the substitutions that were found in each terminal branch, between 10 and 13\% were $\Sphy^{\text{b}}$ (table~S1).
    In principle, beneficial back-mutations are bound to reach fixation more often than neutral substitutions.
    Hence, we calculated the $\dnds$ ratio of non-synonymous over synonymous divergence for all terminal lineages, restricting the non-synonymous changes to our set of $\Sphy^{\text{b}}$.
    We obtained values between 1.15 and 1.68 in the different lineages (table~S1), implying that $\Sphy^{\text{b}}$ reached fixation more frequently than synonymous mutations that are supposed to be neutral, which means that these back-mutations are effectively beneficial.

@@

    \subsection*{Selection in populations}
    Second, we assessed whether our scaled selection coefficients ($\Sphy$) were predictive of selective forces exerted at the population level.
    To achieve this, we reconstructed the ancestral genome of each population to polarize single nucleotide polymorphism (SNPS) segregating in the population.
    The scaled selection coefficient ($\Sphy$) predicted for every non-synonymous SNPs allow us to classify them as either $\Sphy^{\text{d}}$, $\Sphy^{\text{n}}$ or $\Sphy^{\text{b}}$ (fig.~\ref{fig:method}B-C).

    In humans, some SNPs have been associated to specific clinical prognosis terms\cite{landrum_clinvar_2018}. Those clinical terms have been obtained by clinical evaluation of the impact of variants on human Mendelian disorders\cite{landrum_clinvar_2018}. Thus they provide an orthogonal estimate of the effect of a mutation on human health, giving us an opportunity to evaluate whether this effect is in accordance with its predicted fitness effect.
    To do so, we investigated whether the non-synonymous SNPs classified as $\Sphy^{\text{d}}$ or as $\Sphy^{\text{b}}$ showed an enrichment in specific clinical terms in comparison to SNPs classified as $\Sphy^{\text{n}}$.
    Our results show that SNPs predicted as deleterious are associated to clinical terms such as \textit{Likely Pathogenic} and \textit{Pathogenic}, implying that, in general, the selective pressure of a mutation exerted across mammals is also predictive of its clinical effect in humans (table~S6).
    Conversely, back-mutations are associated with clinical terms such as \textit{Benign} and \textit{Likely Benign}, which shows that back-mutations are less likely to be functionally damaging.

    In addition to clinical prognosis, frequencies at which SNPs are segregating within populations also provide information on their selective effects.
    For instance, deleterious SNPs are usually segregating at lower frequencies because of the effect of purifying selection, which tends to remove them from the population.
    By gathering information across many SNPs, it is possible to estimate the distribution of fitness effects (DFE) at the population scale, taking synonymous SNPs as a neutral expectation\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016, tataru_inference_2017}.
    From the estimated DFE, we can derive the proportion of beneficial mutations ($\PpolyAdv$), nearly-neutral mutations ($\PpolyNeutral$) and deleterious mutations ($\PpolyDel$) at the population scale.
    These approaches offer a unique opportunity to contrast selection coefficients estimated at the population scale ($\Spop$) and at the phylogenetic scale ($\Sphy$).

    Across our classes of selection ($\Sphy^{\text{d}}$, $\Sphy^{\text{n}}$ and $\Sphy^{\text{b}}$), one can ultimately estimate the proportion of false positives and false negatives, leading to an estimation of \textit{precision} and \textit{recall} (see Materials \& Methods).
    Across 28 populations of different mammal species, mutations predicted to be deleterious at the phylogenetic scale were indeed purged at the population scale, with a \textit{precision} in the range of 89--97\% (table.~\ref{table:proba}).
    Conversely, a \textit{recall} in the range of 95--98\% implies that mutations found to be deleterious at the population scale were most likely predicted to be so at the phylogenetic scale (table.~\ref{table:proba}).
    Altogether, purifying selection is largely predictable and amino acids with negative fitness effects across mammals were effectively purged away in each population.

    Mutations predicted as $\Sphy^{\text{n}}$ were effectively composed of a mix of neutral and selected mutations with varying \textit{precision} (18--74\%) and \textit{recall} (21--52\%) across the different populations (table.~\ref{table:proba}).
    The variable proportions found between populations can be explained by the effective number of individuals in the population ($\Ne$), which is known to be a major driver of selection efficacy.
    Since $\Ne$ is not directly measurable, we used Watterson's $\theta_w$ as a proxy to order populations by their levels of synonymous diversity.
    Higher synonymous diversity was associated with a smaller proportion of nearly-neutral mutations (fig.~\ref{fig:diversity}A).
    This result follows the prediction of the nearly-neutral theory and suggests that in populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}), discrimination between beneficial and deleterious mutations is more likely to occur (fig~S1-S9).
    Conversely, in populations with lower diversity (e.g.~\textit{Homo}) many more mutations are effectively neutral.

    %Of note, mutations with $\Sphy>1$ only represented ca.~1.5\% of all possible new mutations, but ca.~8.5\% of all observed mutations in the population (blue in fig.~\ref{fig:homo-afr-results}A-B, table~S3).
    Finally, mutations predicted to be beneficial back-mutations based on the selection coefficients at the phylogenetic scale ($\Sphy^{\text{b}}$) were indeed beneficial for individuals currently bearing them, with a precision in the range of 15--52\% (table.~\ref{table:proba}).
    This result confirms that selection towards amino acids that are restoring existing functions is ongoing in these populations.
    Importantly, the \textit{recall} value in this case, computed as $\proba [ \Sphy^{\text{b}}  \given  \Spop^{\text{b}}]$, is the probability for a beneficial mutation at the population scale to be a beneficial back-mutation, that is, going toward a fitter amino-acid (fig.~\ref{fig:method}F).
    In other words, the \textit{recall} value quantifies the amount of beneficial mutations that are restoring damaged genomes instead of creating adaptive innovations.
    Across the 28 populations, this proportion is in the range of 11--70\% (table.~\ref{table:proba}).

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.homo-afr-results}
        \caption{
            (A) Distribution of fitness effects ($\Sphy$), predicted for all possible mutations away from the ancestral human genome.
            Mutations are divided into three classes of selection: deleterious ($\Sphy^{\text{d}}$ in red), nearly-neutral ($\Sphy^{\text{n}}$ in green) and beneficial ($\Sphy^{\text{b}}$ blue, supposedly beneficial back-mutations).
            (B) Distribution of fitness effects ($\Sphy$) for all observed substitutions along the \textit{Homo} branch after the \textit{Homo}-\textit{Pan} splits.
            If there are less substitutions observed than expected, this class is thus undergoing purifying selection, as is the case of for $\Sphy^{\text{d}}$.
            (C) The site-frequency spectrum (SFS) in humans of African descent for a random sample of 16 alleles (mean in solid line and standard deviation in color shade) for each class of selection and for synonymous mutations, which are supposedly neutral (black). The SFS represents the proportion of mutations (y-axis) with a given number of derived alleles in the population (x-axis).
            At high frequencies, deleterious mutations are underrepresented.
            (D) Proportion of beneficial $\PpolyAdv$, nearly-neutral $\PpolyNeutral$ and deleterious mutations $\PpolyDel$ estimated at the population scale for each class of selection at the phylogenetic scale. Proportions depicted here are not weighted by mutational opportunities.
        }
        \label{fig:homo-afr-results}
    \end{figure*}

>>>>

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\der}{\text{d}}
\newcommand{\e}{\text{e}}
\newcommand{\Ne}{N_{\text{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\text{fix}}}
\newcommand{\dn}{d_N}
\newcommand{\ds}{d_S}
\newcommand{\dnds}{\dn / \ds}
\newcommand{\Sphy}{S_{0}}
\newcommand{\SphyDel}{\Sphy^{d}}
\newcommand{\SphyNeu}{\Sphy^{n}}
\newcommand{\SphyBen}{\Sphy^{b}}
\newcommand{\Sphyclass}{x}
\newcommand{\SphyclassAlt}{y}
\newcommand{\given}{\mid}
\newcommand{\Spop}{S}
\newcommand{\SpopDel}{\Spop^{d}}
\newcommand{\SpopNeu}{\Spop^{n}}
\newcommand{\SpopBen}{\Spop^{b}}
\newcommand{\ProbaPopDel}{\proba [ \SpopDel]}
\newcommand{\ProbaPopNeu}{\proba [ \SpopNeu ]}
\newcommand{\ProbaPopBen}{\proba [ \SpopBen ]}
\newcommand{\AdvMean}{\beta_b}
\newcommand{\DelMean}{\beta_d}
\renewcommand{\baselinestretch}{1.5}
\renewcommand{\arraystretch}{0.6}
\linenumbers
\frenchspacing

@@

    Adaptation is one of the main processes shaping the diversity of forms and functions across the tree of life~\cite{darwin_origin_1859}.
    Evolutionary adaptation is tightly linked to environmental change and species responding to this change~\cite{merrell_adaptive_1994, gavrilets_adaptive_2009}.
    For adaptation to occur, there must be variation within populations, which mostly appears via mutations in the DNA sequence.
    While neutral mutations will not impact an individual's fitness, deleterious mutations have a negative effect, and beneficial mutations improve their bearer's fitness.
    A beneficial mutation is thus more likely than a neutral mutation to invade the population (fixation), resulting in a substitution at the species level.
    Upon environmental change, because adaptive beneficial mutations toward new fitness optima are more likely, the number of substitutions also increases (fig.~\ref{fig:fitness-landscape}A).
    An increased substitution rate is thus commonly interpreted as a sign of adaptation~\cite{mcdonald_adaptative_1991, smith_adaptive_2002, welch_estimating_2006}.
    The availability of large-scale genomic data and the development of theoretical models have enabled the detection and quantification of substitution rate increases across genes and lineages~\cite{yang_statistical_2000, eyre-walker_genomic_2006, moutinho_variation_2019}.
    These approaches, now common  practice in evolutionary biology, helped better understand the processes underpinning the rates of molecular evolution.
    However, a collateral effect has been conflating beneficial mutations with adaptive evolution when adaptive evolution is not the only process that can lead to beneficial mutations~\cite{charlesworth_other_2007, mustonen_fitness_2009}.

    In a constant environment, a deleterious mutation can reach fixation by genetic drift~\cite{Ohta1992}.
    A new mutation restoring the ancestral fitness will thus be beneficial (fig.~\ref{fig:fitness-landscape}B), even though the environment has not changed~\cite{hartl_compensatory_1996, sella_application_2005, mustonen_fitness_2009, cvijovic_fate_2015}.
    The restoration of the ancestral fitness can either happen through a mutation at a different locus – called a compensatory mutation~\cite{hartl_compensatory_1996, mustonen_fitness_2009}, or at the locus of the initial mutation – called a beneficial back-mutation~\cite{piganeau_estimating_2003, charlesworth_other_2007}.
    While compensatory mutations change the sequence and thus induce genetic diversification, beneficial back-mutations reduce the genetic diversity and do not contribute to genetic innovation.
    Although Tomoko Ohta considered beneficial back-mutations negligible in her nearly-neutral theory~\cite{Ohta1992}, their importance has now been acknowledged for expanding populations~\cite{charlesworth_other_2007}.
    However, differentiating between an adaptive mutation and a beneficial back-mutation remains challenging\cite{chi_detecting_2020}.
    In this study, we seek empirical evidence for beneficial back-mutations and test long-standing related questions, such as how common they are, and what the probability is for any beneficial mutation observed in current populations to be a back-mutation.
    By addressing these questions, we provide novel insight into the process of adaptation and, more broadly, into how natural selection has affected protein-coding DNA sequences throughout evolution.

\section*{Results}
    \subsection*{Fitness landscape reconstruction}

    An adaptive mutation responding to a change in the environment or a beneficial back-mutation have equivalent consequences for their bearer~\cite{charlesworth_other_2007}.
    Similarly, at the level of the population, both types of mutations will result in a positive transmission bias of the beneficial allele.
    However, at the macro-evolutionary scale the consequences of these two types of mutations are fundamentally different.
    While adaptive mutations promote phenotype diversification (fig.~\ref{fig:fitness-landscape}C), beneficial back-mutations promote phenotype stability and may help preserve well established biological systems (fig.~\ref{fig:fitness-landscape}D).
    Additionally, the direction of adaptive evolution is unpredictable because it is caused by an unpredictable change in the environment and, hence, in the underlying fitness landscape~\cite{bazykin_changing_2015}.
    On the other hand, beneficial back-mutations are predictable if the underlying stable fitness landscape is known because changes from non-optimal amino acids to optimal ones will be expected.
    In this context, the mutation-selection framework permits to relate the patterns of substitution along a phylogeny to the underlying fitness landscape~\cite{halpern_evolutionary_1998, mccandlish_modeling_2014}.
    Such mutation-selection models applied to protein coding DNA sequences allow us to estimate relative fitnesses for all amino-acids for each site of the sequence, explicitly assuming that the underlying fitness landscape is stable along the phylogenetic tree~\cite{rodrigue_mechanistic_2010, tamuri_estimating_2012, rodrigue_detecting_2017}.
    Importantly, because mutation-selection models at the phylogenetic scale are based on population-genetics equations, their estimates of fitness effects are directly interpretable as fitness effects for an individual within a population.
    %These methods account for phylogenetic relationships, the structure of the genetic code and accommodate for underlying mutational biases~\cite{rodrigue_mechanistic_2010, tamuri_estimating_2012, latrille_improved_2022}.

    Theoretically, we can thus use large-scale genomic data to assess whether the fitnesses estimated at the phylogenetic scale are predictive of the fitness effect at the population scale, for both deleterious and beneficial back-mutations.
    The placental mammals represent an excellent study system to perform such analysis.
    Having originated $\sim$100 Mya~\cite{kumar_timetree_2017}, they diversified quickly.
    Additionally, polymorphism data are available for many species~\cite{howe_ensembl_2021} as high quality protein-coding DNA alignments across the genome~\cite{ranwez_orthomam_2007, scornavacca_orthomam_2019}.
    By restricting our analysis to protein-coding orthologous genes across all the phylogeny, we restrict our analysis to genes mostly under functional constrains across mammals.
    For each gene, fitting the mutation-selection model to the multi-species sequence alignment assuming that the underlying fitness landscape is stable along the phylogenetic tree allows obtain relative fitnesses for all amino-acids for each site of the alignment (fig~\ref{fig:method}A).
    From the fitness landscape, we can calculate the scaled selection coefficient $\Sphy$ for any possible amino-acid change at each position.
    From the selection coefficient, we can classify any mutation as either a deleterious mutation toward a less fit amino-acid ($\SphyDel$), a nearly-neutral mutation ($\SphyNeu$) or a beneficial back-mutation toward a known fitter amino-acid ($\SphyBen$).
    Having identified which potential DNA changes represent beneficial back-mutations (fig~\ref{fig:method}B), we retrieved polymorphism data from 28 populations belonging to 6 genera (\textit{Equus}, \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus} and \textit{Homo}) to assess the presence of beneficial back-mutations at the population scale.
    We focused on currently segregating mutations within populations and also on substitutions in the terminal branches, and checked if any of these observed changes where beneficial back-mutations (fig~\ref{fig:method}C-E).
    Essentially, by integrating large scale genomic datasets at both phylogenetic and population scales we have been able to estimate the amount of beneficial back-mutations across the entire exome for 6 different mammalian genera (fig~\ref{fig:method}F-H).

@@

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.method.proba}
        \caption{
            Selection coefficients at the phylogenetic and population scales.
            At the phylogenetic scale (A), the amino-acid Wrightian fitness for each site is estimated from protein-coding DNA alignments using mutation-selection codon models.
            For every possible mutation, the difference in amino-acid fitness before and after the mutation allows us to compute the selection coefficient at the phylogenetic scale ($\Sphy$).
            Mutations with $\Sphy < -1$ are considered deleterious ($\SphyDel$ in red), and mutations with $-1 < \Sphy < 1$ are considered nearly-neutral ($\SphyNeu$ in green).
            Mutations with $\Sphy > 1$ are toward a fitter amino-acids, and thus are considered beneficial back-mutations ($\SphyBen$ in blue), repairing existing functions (B).
            At the population scale (C), each observed single nucleotide polymorphism (SNP) segregating in the population is classified according to its $\Sphy$ value.
            Similarly, substitutions along the terminal branch are classified depending on their $\Sphy$ value.
            Together, substitutions and polymorphisms are contrasted to synonymous changes used as a neutral expectation such as estimate selection coefficients at the population scale ($\Spop$).
            For each class of selection ($\SphyBen$, $\SphyNeu$, $\SphyDel$), we can thus assess whether $\Sphy$ predicts $\Spop$ (E), and compute \textit{precision} (F-G) and \textit{recall} (F-H) for each class.
            The \textit{recall} value for the class $\SphyBen$ is the probability of back-mutations among all beneficial ones (H).
        }
        \label{fig:method}
    \end{figure*}

    \subsection*{Selection along the terminal branches}
    First, we assessed whether fitness effects derived from the mutation-selection model at the phylogenetic scale are predictive of selection occurring in terminal branches.
    To achieve this, we recovered the mutations that reached fixation in the terminal branches of 6 different mammalian genera.
    Only mutations fixed in a population were considered as substitutions in the corresponding branch, while segregating mutations were discarded from this analysis.
    For each substitution identified in the terminal branches, we obtained its scaled selection coefficient ($\Sphy$) as the change in amino-acid fitness at the corresponding codon site in the mammalian alignment (fig.~\ref{fig:method}A-C).
    Importantly, the mammalian alignment used to estimate amino-acid fitness landscape does not include the genera for which we reconstructed the substitutions, to avoid fallacy and circularity in the reasoning.
    Because $\Sphy$ was based on estimation of fitness effects at the mammalian scale, mutations with $\Sphy>1$ were considered beneficial back-mutations ($\SphyBen$), toward a fitter amino acid.
    Among all the substitutions that were found in each terminal branch, between 10 and 13\% were $\SphyBen$ (table~S1).
    In principle, beneficial back-mutations are bound to reach fixation more often than neutral substitutions.
    Hence, we calculated the $\dnds$ ratio of non-synonymous over synonymous divergence for all terminal lineages, restricting the non-synonymous changes to our set of $\SphyBen$.
    We obtained values between 1.15 and 1.68 in the different lineages (table~S1), implying that $\SphyBen$ reached fixation more frequently than synonymous mutations that are supposed to be neutral, which means that these back-mutations are effectively beneficial.

@@

    \subsection*{Selection in populations}
    Second, we assessed whether our scaled selection coefficients ($\Sphy$) were predictive of selective forces exerted at the population level.
    To achieve this, we reconstructed the ancestral genome of each population to polarize single nucleotide polymorphism (SNPS) segregating in the population.
    The scaled selection coefficient ($\Sphy$) predicted for every non-synonymous SNPs allow us to classify them as either $\SphyDel$, $\SphyNeu$ or $\SphyBen$ (fig.~\ref{fig:method}B-C).

    In humans, some SNPs have been associated to specific clinical prognosis terms~\cite{landrum_clinvar_2018}. Those clinical terms have been obtained by clinical evaluation of the impact of variants on human Mendelian disorders~\cite{landrum_clinvar_2018}. Thus they provide an orthogonal estimate of the effect of a mutation on human health, giving us an opportunity to evaluate whether this effect is in accordance with its predicted fitness effect.
    To do so, we investigated whether the non-synonymous SNPs classified as $\SphyDel$ or as $\SphyBen$ showed an enrichment in specific clinical terms in comparison to SNPs classified as $\SphyNeu$.
    Our results show that SNPs predicted as deleterious are associated to clinical terms such as \textit{Likely Pathogenic} and \textit{Pathogenic}, implying that, in general, the selective pressure of a mutation exerted across mammals is also predictive of its clinical effect in humans (table~S6).
    Conversely, back-mutations are associated with clinical terms such as \textit{Benign} and \textit{Likely Benign}, which shows that back-mutations are less likely to be functionally damaging.

    In addition to clinical prognosis, frequencies at which SNPs are segregating within populations also provide information on their selective effects.
    For instance, deleterious SNPs are usually segregating at lower frequencies because of the effect of purifying selection, which tends to remove them from the population.
    By gathering information across many SNPs, it is possible to estimate the distribution of fitness effects (DFE) at the population scale, taking synonymous SNPs as a neutral expectation~\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016, tataru_inference_2017}.
    From the estimated DFE, we can derive the proportion of beneficial mutations ($\ProbaPopBen$), nearly-neutral mutations ($\ProbaPopNeu$) and deleterious mutations ($\ProbaPopDel$) at the population scale.
    These approaches offer a unique opportunity to contrast selection coefficients estimated at the population scale ($\Spop$) and at the phylogenetic scale ($\Sphy$).

    Across our classes of selection ($\SphyDel$, $\SphyNeu$ and $\SphyBen$), one can ultimately estimate the proportion of false positives and false negatives, leading to an estimation of \textit{precision} and \textit{recall} (see Materials \& Methods).
    Across 28 populations of different mammal species, mutations predicted to be deleterious at the phylogenetic scale were indeed purged at the population scale, with a \textit{precision} in the range of 89--97\% (table.~\ref{table:proba}).
    Conversely, a \textit{recall} in the range of 95--98\% implies that mutations found to be deleterious at the population scale were most likely predicted to be so at the phylogenetic scale (table.~\ref{table:proba}).
    Altogether, purifying selection is largely predictable and amino acids with negative fitness effects across mammals were effectively purged away in each population.

    Mutations predicted as $\SphyNeu$ were effectively composed of a mix of neutral and selected mutations with varying \textit{precision} (18--74\%) and \textit{recall} (21--52\%) across the different populations (table.~\ref{table:proba}).
    The variable proportions found between populations can be explained by the effective number of individuals in the population ($\Ne$), which is known to be a major driver of selection efficacy.
    Since $\Ne$ is not directly measurable, we used Watterson's $\theta_w$ as a proxy to order populations by their levels of synonymous diversity.
    Higher synonymous diversity was associated with a smaller proportion of nearly-neutral mutations (fig.~\ref{fig:diversity}A).
    This result follows the prediction of the nearly-neutral theory and suggests that in populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}), discrimination between beneficial and deleterious mutations is more likely to occur (fig~S1-S9).
    Conversely, in populations with lower diversity (e.g.~\textit{Homo}) many more mutations are effectively neutral.

    %Of note, mutations with $\Sphy>1$ only represented ca.~1.5\% of all possible new mutations, but ca.~8.5\% of all observed mutations in the population (blue in fig.~\ref{fig:homo-afr-results}A-B, table~S3).
    Finally, mutations predicted to be beneficial back-mutations based on the selection coefficients at the phylogenetic scale ($\SphyBen$) were indeed beneficial for individuals currently bearing them, with a precision in the range of 15--52\% (table.~\ref{table:proba}).
    This result confirms that selection towards amino acids that are restoring existing functions is ongoing in these populations.
    Importantly, the \textit{recall} value in this case, computed as $\proba [ \SphyBen  \given  \SpopBen]$, is the probability for a beneficial mutation at the population scale to be a beneficial back-mutation, that is, going toward a fitter amino-acid (fig.~\ref{fig:method}F).
    In other words, the \textit{recall} value quantifies the amount of beneficial mutations that are restoring damaged genomes instead of creating adaptive innovations.
    Across the 28 populations, this proportion is in the range of 11--70\% (table.~\ref{table:proba}).

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.homo-afr-results}
        \caption{
            (A) Distribution of fitness effects ($\Sphy$), predicted for all possible mutations away from the ancestral human genome.
            Mutations are divided into three classes of selection: deleterious ($\SphyDel$ in red), nearly-neutral ($\SphyNeu$ in green) and beneficial ($\SphyBen$ blue, supposedly beneficial back-mutations).
            (B) Distribution of fitness effects ($\Sphy$) for all observed substitutions along the \textit{Homo} branch after the \textit{Homo}-\textit{Pan} splits.
            If there are less substitutions observed than expected, this class is thus undergoing purifying selection, as is the case of for $\SphyDel$.
            (C) The site-frequency spectrum (SFS) in humans of African descent for a random sample of 16 alleles (mean in solid line and standard deviation in color shade) for each class of selection and for synonymous mutations, which are supposedly neutral (black). The SFS represents the proportion of mutations (y-axis) with a given number of derived alleles in the population (x-axis).
            At high frequencies, deleterious mutations are underrepresented.
            (D) Proportion of beneficial $\ProbaPopBen$, nearly-neutral $\ProbaPopNeu$ and deleterious mutations $\ProbaPopDel$ estimated at the population scale for each class of selection at the phylogenetic scale. Proportions depicted here are not weighted by mutational opportunities.
        }
        \label{fig:homo-afr-results}
    \end{figure*}
